\chapter{Introduction}

\section{Project Repository}
The source code for this project is available at \url{https://github.com/ArielCNunes/TRiM-FYP}. The repository is organised into three top-level directories:

\begin{itemize}                                                                                                              
    \item \textbf{\texttt{backend/}}: Contains the Spring Boot REST API, structured as a standard Maven project. Key packages include \texttt{controller}, \texttt{service}, \texttt{repository}, \texttt{entity}, and \texttt{tenant}, reflecting the layered architecture and multi-tenant logic at the core of the system. Supporting packages handle payment processing, email notifications, booking reminders, and test data seeding.                                                                                            
    \item \textbf{\texttt{frontend/}}: A React application built with TypeScript and Vite, styled with Tailwind CSS. The source is organised into \texttt{pages}, \texttt{components}, \texttt{features}, \texttt{hooks}, and \texttt{store} directories, separating UI concerns from state management and API integration.
    \item \textbf{\texttt{docs/}}: Contains this dissertation in \LaTeX{} format, along with records of client and supervisor meetings maintained throughout the project.
\end{itemize}  

\section{Context}
Small businesses such as barber shops that rely solely on walk-in customers face several operational challenges: limited business insights, customer friction when services are unavailable at their preferred time, and the burden of manual bookkeeping. Digitalising these operations bridges the gap between the business and its customers, leading to a more established market presence and streamlined day-to-day management.

\textit{TRiM} is a platform designed to support the full range of barber shop operations, from customer bookings and payment processing to employee management and business analytics. Rather than serving a single business, \textit{TRiM} is built to allow multiple businesses to operate on the platform independently, each with its own isolated data and user base. One instance of the software serves multiple businesses, this is known as multi-tenant architecture, and it is the core focus of this dissertation.

\section{Project Objectives}
The objective of this project is to deliver a real and robust software solution to local service-based businesses. The following goals have been defined:
\begin{itemize}
    \item \textbf{Customer Booking System}: Enable customers to book services online, select their preferred barber, choose from categorized services, and pick available time slots while the system prevents double bookings and scheduling conflicts.
    
    \item \textbf{Payment Integration}: Integrate Stripe for secure online payment processing, supporting both full online payments and deposit-based bookings where customers pay a percentage upfront and settle the outstanding balance in-shop, driven by real client requirements.
    
    \item \textbf{Business Administration}: Provide business owners with a comprehensive admin dashboard to manage service categories, individual services with pricing and duration, and employee (barber) profiles. Include analytics such as total revenue, booking statistics, and popular services.
    
    \item \textbf{Barber Management}: Allow barbers to manage their own availability by setting weekly working hours, defining break periods, and viewing/managing their upcoming bookings with the ability to confirm, complete, or handle cancellations and no-shows.
    
    \item \textbf{Customer Management}: Enable administrators to view customer history, track no-show counts, and blacklist problematic customers to protect business operations.
    
    \item \textbf{Multi-Tenant Architecture}: Design the platform to support multiple independent businesses, each with their own subdomain (business slug), isolated data, and separate admin/barber/customer user bases.
    
    \item \textbf{Role-Based Access Control}: Implement distinct user roles (CUSTOMER, BARBER, ADMIN) with appropriate permissions and tailored interfaces for each role's specific needs.
\end{itemize}

\section{Dissertation Focus}
This dissertation focuses primarily on the architectural decisions behind \textit{TRiM}, particularly the database design patterns and data isolation techniques employed. Through \textit{TRiM}, we explore how to build a multi-tenant software application that is both scalable and robust, leveraging efficient design patterns that allow the system to grow without significant cost or performance overhead. We also investigate how multiple businesses can share the same platform whilst maintaining strong performance guarantees and strict security boundaries.

\section{Success Metrics}
From a functional standpoint, the system must handle all core operations reliably: bookings should complete without conflicts, payments should process securely, and the admin dashboard should provide accurate insights into business operations.

Beyond functionality, performance represents a critical area of evaluation. The system will be benchmarked under various conditions to assess its resilience. This includes measuring response times for key API endpoints, evaluating database performance as the number of tenants and bookings increases, and verifying that heavy usage by one business does not degrade the experience for others. Tenant isolation extends beyond security considerations, it also ensures fair distribution of system resources across all users.

Security constitutes another essential metric. With multiple businesses and their customer data at stake, the system must demonstrate robust protection mechanisms. Cross-tenant data leakage would represent a critical failure, necessitating thorough testing of all isolation mechanisms to prevent such occurrences.

Finally, usability remains a key consideration. The primary objective of \textit{TRiM} is to streamline operations for business owners and their customers. An unintuitive interface or cumbersome booking flow would undermine the value of the underlying architecture.

\section{Chapters Overview}

\subsection*{Chapter 2: Methodology}
Outlines the development approach used. It describes the agile workflow with a Kanban board in GitHub Projects, the user stories used to gather requirements, and the feedback loop with weekly supervisor meetings and regular client demonstrations with V7 Barbers. The chapter also covers the tools used for development, the role of generative AI, testing, and the evaluation strategy built around Gatling load testing and seeded test data.

\subsection*{Chapter 3: Technology Review}
Looks at the three primary multi-tenant database architectures: separate databases, separate schemas, and shared schemas. Analyses their trade-offs in terms of isolation, cost, and scalability. The chapter surveys data isolation techniques including filter-based and permission-based approaches such as PostgreSQL Row-Level Security, and reviews related work on adaptive schema design, and recent benchmarking frameworks. It identifies the gap that motivates this project: a lack of empirical evidence on combining shared-schema designs with database-level security enforcement.

\subsection*{Chapter 4: System Design}
Presents the architecture of \textit{TRiM}, from the RESTful API and entity-relationship model to the multi-tenant implementation built on subdomain-based routing, tenant-aware schema design, and Row-Level Security. It details core operational flows including booking and payment processing with Stripe, race condition prevention, JWT-based authentication with role-based access control, and cross-subdomain token exchange. The chapter also covers the React frontend architecture and the Docker-based deployment strategy.

\subsection*{Chapter 5: System Evaluation}
Evaluates the system against the success metrics defined in this chapter. Performance is assessed through Gatling load tests that measure API response times and database behaviour as tenant and booking volumes increase. Security is verified through tenant isolation testing to confirm that no cross-tenant data leakage occurs. The chapter concludes with a critical analysis of the system's strengths and the limitations of the current design.

\subsection*{Chapter 6: Conclusion}
Reflects on the extent to which the project objectives were met, sums up the key findings from the system evaluation, and discusses the broader implications of the architectural choices made. It concludes with directions for future work, including potential improvements to the platform and areas where the multi-tenant design could be extended.
